### 2022-05-06 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@21da99f6e7268f26f8e019dd4c4ae5cb5f973aec 🚀

Spillover risk is not unidimensional. From the standpoint of an animal
community, \emph{i.e.} a pool of suitable hosts, it is driven by a
multiplicity of factors (Plowright et al. 2017). The global richness of
hosts is one such component commonly mentioned/analysed (see \emph{e.g.}
Anthony et al. 2017 for coronaviruses), but there is an argument to be
made that species who are not competent (or know) hosts of a specific
virus genus may not factor into this (Plowright et al. 2015), or that
species who are assumed to share viruses at different rates should be
weighted accordingly (Albery et al. 2020). In mammals, key functional
traits (for which phylogeny is a reasonable proxy) are determinants of
the spillover potential (Olival et al. 2017); these include, notably,
body mass and affinity for urban environments (Albery et al. 2022).
Finally, especially when the pool of potential hosts spans the entire
globe, there may be local host pools that are highly unique; not having
been observed in other locations, these can act on the overall risk
either by providing novel contact opportunities, reflecting unique
host-environment combinations (Engering et al. 2013), or facilitating
rapid evolutionary changes in specialism of their pathogens (Agosta et
al. 2010). In the specific case of generalist pathogens, there is
conceptual and empirical support to the idea that these community- level
mechanisms are even more important in driving the overall risk (Power
and Mitchell 2004).

Bats are important reservoir hosts for different classes of
microorganisms (Chu 2008, Donaldson 2010, Li 2010), some of which can
threaten human health. Chiropterans emerged around 64 million years ago
and are one of the most diverse mammalian orders, with an estimated
richness of more than 14000 species (Peixoto et al. 2018, Simmons and
Cirranello 2020). They exhibit a broad variety of habitat use,
behaviour, and feeding strategies, resulting in their playing an
essential role in the delivery of several ecosystem services tied to
important ecosystem-derived benefits (Kasso and Balakrishnan 2013). For
example, over two-thirds of bats are know to be either obligate or
facultative insectivorous mammals, therefore playing an important role
in the regulation of insect pests that can affect crops
(Williams-Guillén et al. 2008, Voigt and Kingston 2016), and vectors of
diseases that put a risk on human health (Gonsalves et al. 2013a, b).
Because bats are globally distributed and have a long evolutionary
history, phylogeographic and biogeographic approaches are required to
shed light on the extant distribution of coevolutionary processes
between bats and the pathogens they carry. Not all areas in which bats,
viruses, and human are co-occuring are facing a risk of spillover
towards human populations, and the areas in which this risk exist may
not be facing risks of the same nature and magnitude.

In this paper, we examine the biogeographic structure of
bats-betacoronaviruses associations, based on a curated dataset of known
and recently discovered hosts. This work is important both as a
description of the bats-betacoronavirus complex, but aslo because more
broadly, bats are known reservoirs for a variety of emerging viruses and
pathogens (Calisher et al. 2006, Melaun et al. 2014), making balancing
the needs for bat conservation and disease prevention a potentially
difficult act and a source of human-wildlife conflicts, especially in
more densely populated areas (Stone et al. 2015, Rego et al. 2015). By
drawing on concepts from the Geographic Mosaic Theory of Coevolution
(Thompson 2005), we turn these associations into a spatially explicit
additive mapping of zoonotic risk components, which reveals extreme
heterogeneity of risk at the global scale; furthermore, we identify the
Amazon and South-Eastern Asia as hotspots of phylogenetic
distinctiveness of betacoronaviruses (Anthony et al. 2017);
surprisingly, current data suggest that viral sharing between hosts in
high in the Amazon and low in South-Eastern Asia, which has the
potential to result in different evolutionary dynamics between these two
regions.

>>>>

Spillover risk is multidimensional and complicated. From the standpoint
of an animal community, \emph{i.e.} a pool of suitable hosts, it is
driven by a multiplicity of factors (Plowright et al. 2017). Proxies for
the global richness of hosts is one such component commonly analysed
(see \emph{e.g.} Anthony et al. 2017 for coronaviruses), but there is an
argument to be made that species who are not competent (or known) hosts
of a specific virus genus may not factor into this (Plowright et al.
2015); this is especially true as competence data increases predictive
power when the taxonomic scope of hosts of a viral family increases
(\textbf{Becker2020InfInt?}, \textbf{Mull2022VirIso?}). Similarly, host
species who are assumed to share viruses at different rates should be
weighted accordingly (Albery et al. 2020). In mammals, key functional
traits (for which phylogeny is a reasonable proxy) are determinants of
the spillover potential (Olival et al. 2017); these include, notably,
body mass and affinity for urban environments (Albery et al. 2022).
Finally, especially when the pool of potential hosts spans the entire
globe, there may be local host pools that are highly unique; not having
been observed in other locations, these can act on the overall risk
either by providing novel contact opportunities, reflecting unique
host-environment combinations (Engering et al. 2013), or facilitating
rapid evolutionary changes in specialism of their pathogens (Agosta et
al. 2010). In the specific case of generalist pathogens (which is the
case many viruses in the betacoronavirus genus, see \emph{e.g.}
\textbf{MacLean2021NatSel?}), there is conceptual and empirical support
to the idea that these community-level mechanisms are even more
important in driving the overall risk (Power and Mitchell 2004).

Bats are important reservoir hosts for different classes of
microorganisms (Chu 2008, Donaldson 2010, Li 2010), many of which can
threaten human health (\textbf{Letko2020BatVir?},
\textbf{VanBrussel2022ZooDis?}). Chiropterans emerged around 64 million
years ago and are one of the most diverse mammalian orders, with an
estimated richness of more than 1400 species (Peixoto et al. 2018,
Simmons and Cirranello 2020). They exhibit a broad variety of habitat
use, behaviour, and feeding strategies, putting them at key positions in
the delivery and provisioning of several ecosystem services, tied to
important ecosystem-derived benefits to human (Kasso and Balakrishnan
2013). For example, bats are an essential component of many
seed-dispersal networks (\textbf{Mello2011MisPar?}). Over two-thirds of
bats are know to be either obligate or facultative insectivores,
therefore playing an important role in the regulation of insect pests
that can affect crops (Williams-Guillén et al. 2008, Voigt and Kingston
2016), and vectors of pathogens that put a risk on human health
(Gonsalves et al. 2013a, b). Because bats are globally distributed and
have a long evolutionary history, phylogeographic and biogeographic
approaches are required to shed light on the contemporary distribution
of coevolutionary processes between bats and the pathogens they host.
Not all areas in which bats, viruses, and human are co-occuring are
facing a risk of spillover towards human populations, and the areas in
which this risk exist may not be facing risks of the same nature and
magnitude.

In this paper, we examine the biogeographic structure of
bats-betacoronaviruses associations, based on a curated dataset of all
confirmed bat hosts of betacoronaviruses. This work is important both as
a description of the bats-betacoronavirus complex, but aslo because more
broadly, bats are known reservoirs for a variety of emerging viruses and
other pathogens (Calisher et al. 2006, Melaun et al. 2014), making
balancing the needs for bat conservation and disease prevention most
likely very difficult and a source of human-wildlife conflicts,
especially in more densely populated areas (Stone et al. 2015, Rego et
al. 2015). By drawing on concepts from the Geographic Mosaic Theory of
Coevolution (Thompson 2005), we turn these associations into a spatially
explicit additive mapping of zoonotic risk components, revealing the
extreme heterogeneity of risk at the global scale. We identify the
Amazon and South-Eastern Asia as hotspots where the phylogenetic
distinctiveness of betacoronaviruses is the highest (Anthony et al.
2017); surprisingly, current data suggest that viral sharing between
hosts is high in the Amazon and low in South-Eastern Asia, which has the
potential to result in different evolutionary dynamics between these two
regions.
